\documentclass{article}
\input{preamble}

\title{Statistics extra questions + references}
\author{Leo Tyrpak}
\date{September 2024}

\begin{document}

\maketitle
I will follow closely and recommend you read chapters 1 and 2 of Stein and Shakarchi measure theory book(https://www.cmat.edu.uy/~mordecki/courses/medida2013/book.pdf).

\section{Sheet 1}
Notes:
\begin{itemize}
    \item Q3: this is another way to think about integrals as linear functionals on the vector space of continuous and bounded functions on \(\R^d\). Essentially a measure \(\mu\) allows you to integrate stuff i.e. functions \(f\) with integral \(\int fd\mu\) and the map \(f\to\int fd\mu\) is a linear functional from the set of continuous functions to \(\R\) and is continuous in the sense that if \(f_n\to f\) then \(\int f_nd\mu\to\int fd\mu\) in a sense I will not make precise. The Riesz-Representation theorem gives you the other direction, that linear functionals on continuous functions are just measures.
\end{itemize}
Extra:
\begin{itemize}
    \item Q1: Calculate \(\limsup\) and \(\liminf\) for the following \(a_n\): \(\cos(n),\cos(n^2)\)
    \item Q1: Prove that if \(g\) continuous and increasing then \(g(\limsup{a_n})=\limsup{g(a_n)}\). Prove similarly for \(\liminf\) and \(g\) continuous and decreasing (are the statements the same in that case). Is it enough for \(g\) to be continuous? Increasing? (If not then find a counterexample).
    \item Q1: prove limit exists iff limsup =liminf of a sequence. How would you define \(\limsup_{t\to a}g(t(\) where \(t\to a\) as convergence of real numbers. Relate this to limsup of sequences.
    \item Q5: does \((c)\) still hold if \(f\) is only assumed to be continuous? Find proof or counterexample. What if \(f\) is only assumed to be differentiable almost everywhere but the derivative is not necessarily continuous.
    \item Do the questions at the end of Chapter of Stein and Shakarchi.
\end{itemize}

\end{document}