\documentclass{article}
\input{preamble}

\title{Numerical analysis extra questions}
\author{Leo Tyrpak}
\date{September 2024}

\begin{document}

\maketitle
These are notes and extra questions for the Part A (second year) course numerical analysis taught in Hilary term (January-March) in Oxford.
\section{Good resources}
I highly recommend An intro to numerical analysis \cite{NAbook}.
Everything I reference will be from \cite{NAbook} unless specified otherwise.
\section{Interpolating polynomials}
\textbf{Reading: }Section 6 covers all the material plus some more on Hermite polynomials.

\textbf{Questions: }I recommend doing all questions at the end of section 6(pg 195).
Although the questions especially relevant for marking would be Q6.4-Q6.9, Q6.10, Q6.12.

\subsection{Numerical integration}
To really see the point of polynomial approximation it is recommended to read Chapter 7 of \cite{NAbook} on numerical integration schemes and do some of the exercises.

\section{LU factorisation}
\textbf{Reading: }I recommend reading through section 2 of \cite{NAbook} to understand LU factorisation.
The course in Oxford mostly covers sections 2.2-2.5.

\textbf{Questions: }As for questions I recommend Q2.1-Q2.6 (pg 82) which follow the course.

\textbf{Computational questions: }It is also a good exercise to implement a function which computes the LU factorisation using your favourite programming language. 
The input would be a square $n\times n$ matrix $A$ and the output is $2$ $n\times n$ matrices $L$ and $U$ which is the $LU$ factorisation.
Once that has been implemented, implement a function a function which takes as input $L$ and $U$ and a vector $b$ and returns $x$ such that $LUx=b$.
\subsection{Condition number}
To learn about condition number, read sections 2.6-2.8 and do Q2.7-2.14 at end of chapter.

\sec
\bibliographystyle{plain}
\bibliography{refs}

\end{document}